Towards Spherical Robots For Mobile Mapping In Human Made Environments\tnoteref{mytitlenote}

### 2021-06-28 Latex/sections/registration.tex
Done reviewer 2 additional points corrections

For the plane extraction the linescans need to be transformed into the project coordinate system.
Various pre-registration methods are suitable for this.
The most simple method is to use the data from an intertial measurement unit for a coarse alignment.
Alternative solutions combine multiple linescans into a metascan and perform registration methods known from terrestrial laser scanning or use a small number of iterations of continuous-time SLAM approaches such as the one presented in~\cite{REMSEN2013}.
This yields maps that can be used for plane registration and hence a point-to-plane correspondences based registration.

@@

After having done the pre-registration, the scans are aligned well enough to make statements about the potential planes in the environment.
To find the planes in the environment, a Randomized Hough transform with an accumulator ball as described in~\cite{3DRESEARCH2011} is used as this method prefers dominant planes such as the building structure over smaller planar surfaces.
In the next subsection, the Hesse-normal form of the plane is required.
For an ideal plane the distance parameter $\rho_{\mathcal{P}_i}$ is computed via $\vec{n}_{\mathcal{P}_i}\cdot\vec{x}_i$, where $\vec{x}_i$ is an arbitrary point on the plane.
To find this point, we first define a convex hull that encloses the plane.
It is defined by the points that lie on the plane with the furthest distance to one another.
We then choose the center point of the convex hull of the plane as $\vec{x}_i$.

\subsection{Point-to-Plane Correspondences}

There are many ways of finding correspondences between points and planes.
However, even with the result not being perfect, we assume already globally registered scans, which is why a simple algorithm is sufficient:
For each plane, consider adding a correspondence for all points closer to the plane than a threshold $\epsilon$.
Depending on the threshold value, some points now have multiple correspondences if their distance to multiple planes is less than $\epsilon$. We then omit all points with correspondences to multiple planes from the optimization to avoid introducing errors due to the plane ambiguity.

@@

Assuming we know now the Hesse-normal form of all planes and all assigned points to these planes, we register the points to have a minimal distance to their respective planes.
The transformation $T(\vec{p}_k)$ of each point $\vec{p}_k$ with respect to a 6 DoF motion is described in homogeneous coordinates using the roll-pitch-yaw ($\varphi-\vartheta-\psi$) Tait-Brian angles as in~\cite{diebel2006representing}. Transforming the result back from homogeneous coordinates and using $C_k$ and $S_k$ to denote the cosine and sine of the angle in the subscript yields:
\begin{align}
	T(\vec{p}_k)  =
    \resizebox{0.35\textwidth}{!}{$\begin{bmatrix}
        x C_\vartheta C_\psi - yC_\varphi S_\psi + z S_\vartheta + t_x\\
        x (C_\varphi S_\psi + C_\psi S_\varphi S_\vartheta) + y(C_\varphi C_\psi - S_\varphi S_\vartheta S_\psi) - z C_\vartheta S_\varphi + t_y\\
        x(S_\varphi S_\psi - C_\varphi C_\psi S_\vartheta) + y(C_\psi S_\varphi + C_\varphi S_\vartheta S_\psi) + z C_\varphi C_\vartheta + t_z
    \end{bmatrix}$}
\end{align}
From this we define the function $D(\varphi,\vartheta,\psi,t_x,t_y,t_z, \vec{p}_k)$ that computes the distance of a point $\vec{p}_k$ to its corresponding plane $\mathcal{P}_i$.
 Omitting the arguments of the function for simplicity:
\begin{align}
\begin{aligned}
    D &= T(\vec{p}_k) \cdot \vec{n}_{\mathcal{P}_i} \\
      &= n_{\mathcal{P}_i}^x (x C_\vartheta C_\psi - yC_\varphi S_\psi + z S_\vartheta + t_x)\\
       &+ n_{\mathcal{P}_i}^y(x (C_\varphi S_\psi + C_\psi S_\varphi S_\vartheta)\\
       &\qquad+ y(C_\varphi C_\psi - S_\varphi S_\vartheta S_\psi) - z C_\vartheta S_\varphi + t_y)\\
       &+ n_{\mathcal{P}_i}^z (x(S_\varphi S_\psi - C_\varphi C_\psi S_\vartheta)\\
       &\qquad+ y(C_\psi S_\varphi + C_\varphi S_\vartheta S_\psi) + z C_\varphi C_\vartheta + t_z)\\
       &- \rho_{\mathcal{P}_i}
\end{aligned}
\end{align}

This distance function is what we want to minimize for all points and their respective planes.
Hence the error function $E$ is chosen as follows:
\begin{align}
	E = \sum_{\forall \mathcal{P}_i}\sum_{\vec{p}_k \in \mathcal{P}_i} \norm[2]{D(\varphi,\vartheta,\psi,t_x,t_y,t_z, \vec{p}_k)}^2
\end{align}
Its gradient follows then immediately:
\begin{align}
	&\nabla E =  \sum_{\forall \mathcal{P}_i}\sum_{\vec{p}_k \in \mathcal{P}_i} \resizebox{0.28\textwidth}{!}{$\begin{bmatrix}\pardiff{\varphi} E&\pardiff{\vartheta} E&\pardiff{\psi} E&\pardiff{t_x} E&\pardiff{t_y} E&\pardiff{t_z} E\end{bmatrix}^T$}\\
    &\Rightarrow\nabla E = \resizebox{0.35\textwidth}{!}{$\sum_{\forall \mathcal{P}_i} \sum_{\vec{p}_k \in \mathcal{P}_i}2D(\vec{\Pi}, \vec{p}_k)\begin{bmatrix}\nabla E_\varphi &\nabla E_\vartheta & \nabla E_\psi & n_{\mathcal{P}_i}^x&n_{\mathcal{P}_i}^y&n_{\mathcal{P}_i}^z\end{bmatrix}^T$}
\end{align}
Where
\begin{align}
    &\begin{alignedat}{3}
        \nabla E_\varphi = & n_{\mathcal{P}_i}^x&&y S_\varphi S_\psi \\
        &+ n_{\mathcal{P}_i}^y&&(x[-S_\varphi S_\psi+C_\varphi C_\psi S_\vartheta] \\&&&+y[-S_\varphi C_\psi -C_\varphi S_\vartheta S\psi] \\&&&- z C_\varphi C_\vartheta)\\
        &+ n_{\mathcal{P}_i}^z&&(x[C_\varphi S_\psi+C_\varphi C_\psi S_\vartheta] \\&&&+y[C_\varphi C_\psi -S_\varphi S_\vartheta S\psi] \\&&&- z S_\varphi C_\vartheta)
    \end{alignedat} \\
    &\begin{alignedat}{3}
        \nabla E_\vartheta = & n_{\mathcal{P}_i}^x(-xS_\vartheta C_\psi + zC_\vartheta) \\
        &+ n_{\mathcal{P}_i}^y(xC_\psi S_\varphi C_\vartheta - yS_\varphi C_\vartheta S_\psi + z S_\vartheta S_\varphi)  \\
        &+ n_{\mathcal{P}_i}^z(-xC_\varphi C_\psi C_\vartheta + yC_\varphi C_\vartheta S_\psi - z C_\varphi S_\vartheta)
    \end{alignedat}\\
    &\begin{alignedat}{3}
       \nabla E_\psi =& n_{\mathcal{P}_i}^x&&\left(-xC_\vartheta S_\psi - yC_\varphi C_\psi\right) \\
       &+ n_{\mathcal{P}_i}^y&&(x[C_\varphi C_\psi - S_\varphi S_\vartheta S_\psi] \\
       &&&+ y[-C_\varphi S_\psi - S_\varphi S_\theta C_\psi]) \\
       &+ n_{\mathcal{P}_i}^z&&(x[S_\varphi C_\psi + C_\varphi S_\psi S_\theta] \\
       &&&+ y[-S_\psi S_ \varphi + C_\varphi S_\vartheta C_\psi]) 
    \end{alignedat}
\end{align}
and $\vec{\Pi}=\begin{bmatrix}\varphi & \vartheta & \psi & t_x & t_y & t_z\end{bmatrix}^T$.

As the gradient is well-defined we minimize the error function with any gradient based method. 
The commonly used, well-known stochastic gradient descent (SDG) algorithm computes 
\begin{align}
    \vec{\Pi}_{k+1} = \vec{\Pi}_{k} - \alpha \nabla E
\end{align}
To accelerate convergence and to improve the found solution further modifications are made.

@@

Another consideration to speed up SDG is to adaptively recalculate $\boldsymbol\alpha$ for each iteration. 
We employ and modify ADADELTA as a technique to do so, which is described in detail in~\cite{zeiler2012adadelta}.
The main idea is the following:
It extends the SDG algorithm by two terms.
First, an exponentially decaying average of past gradients, which is recursively defined as
\begin{align}
    \vec{G}_{k+1} = \rho \vec{G}_{k} + (1 - \rho) {\nabla E}^2
\end{align}
and second, an exponentially decaying average of past changes, which is defined as
\begin{align}
    \vec{X}_{k+1} = \rho \vec{X}_{k} + (1 - \rho) {\boldsymbol\alpha \nabla E}^2
\end{align}
where $\rho \leq 1$ is a decay constant, typically close to $1$.
The root mean squared (RMS) of these quantities are
\begin{align}
    RMS[\vec{G}]_{k} = \sqrt{\vec{G}_{k} + \epsilon}
\end{align}
and 
\begin{align}
    RMS[\vec{X}]_{k} = \sqrt{\vec{X}_{k} + \epsilon}
\end{align}
where $\epsilon > 0$ is a very small constant, typically close to $0$.
It will prevent dividing by zero in the recalculation of $\boldsymbol\alpha$, which is as follows:
\begin{align}
    \boldsymbol\alpha_{k} = \frac{RMS[X]_{k-1}}{RMS[G]_{k}}
    \label{eq:adadeltaalphaupdate}
\end{align} 
For our particular application, ADADELTA behaves a little too aggressively.
Despite giving a good measure on how to adapt $\boldsymbol\alpha$, the algorithm sometimes overshoots, and does not converge.
Therefore, we employ another scaling factor, typically not found in ADADELTA, extending eq.~\eqref{eq:adadeltaalphaupdate} to:
\begin{align}
    \boldsymbol\alpha_{k} = \boldsymbol\alpha_0 \cdot \frac{RMS[X]_{k-1}}{RMS[G]_{k}}
    \label{eq:adafinalized}
\end{align} 
where $\boldsymbol\alpha_0 $ holds the scaling factors for each dimension.

@@

Firstly, we introduce a lock on some dimensions from being optimized by setting the corresponding $\alpha_i$ to zero.
Although 6D optimization generally works, reducing the optimization space is particularly useful if the source of error in the system is known and a model exists.
That way, e.g., a point cloud from a terrestrial laserscanner that experiences irregularities in its spinning mirror is registered by optimizing only over the axis of rotation as no other movement is to be expected.

>>>>

For the plane extraction the linescans need to be transformed into the project coordinate system.
Various pre-registration methods are suitable for this.
The most simple method is to use the data from an intertial measurement unit for a coarse alignment.
Alternative solutions combine multiple linescans into a metascan and perform registration methods known from terrestrial laser scanning \todo{add source here} or use a small number of iterations of continuous-time SLAM approaches such as the one presented in~\cite{REMSEN2013}.
This yields maps that can be used for plane registration and hence a point-to-plane correspondences based registration.

@@

After having done the pre-registration, the scans are aligned well enough to make statements about the potential planes in the environment.
To find the planes in the environment, a Randomized Hough transform with an accumulator ball as described in~\cite{3DRESEARCH2011} is used as this method prefers dominant planes such as the building structure over smaller planar surfaces.
In the next subsection, the Hesse-normal form of the plane is required.
For an ideal plane the orthogonal distance from the origin $\rho_{\mathcal{P}_k}$ is computed via $\vec{n}_{\mathcal{P}_k}\cdot\vec{p}_i$, where $\vec{p}_i = \begin{bmatrix}x_i&y_i&z_i\end{bmatrix}^T$ is an arbitrary point on the plane and $\vec{n}_{\mathcal{P}_k} = \begin{bmatrix}n_{\mathcal{P}_k}^x&n_{\mathcal{P}_k}^y&n_{\mathcal{P}_k}^z\end{bmatrix}^T$ is the normal vector of that plane.
To find this point, we first define a convex hull that encloses the plane.
It is defined by the points that lie on the plane with the furthest distance to one another.
We then choose the center point of the convex hull of the plane as $\vec{p}_i$.

\subsection{Point-to-Plane Correspondences}\label{sec:point-to-plane-correspondences}

There are many ways of solving the segmentation problem of assigning each point to a plane often based on the RANSAC method \cite{Honti2018}, \cite{Gaspers2011} but also deep learning approaches \cite{Engelmann2018}.
The quality of the preregistration directly influences the quality of the plane detection and therefore the quality of the final registration result. 
We assume small enough errors in pre-registration to allow for plane detection.
Which is why a simple algorithm is sufficient:
For each plane, consider adding a correspondence for all points closer to the plane than a threshold $\epsilon$.
Depending on the threshold value, some points now have multiple correspondences if their distance to multiple planes is less than $\epsilon$ \todo{@Fabi add actual $\epsilon$ values to the experiments}. We then omit all points with correspondences to multiple planes from the optimization to avoid introducing errors due to the plane ambiguity.

@@

Assuming we know now the Hesse-normal form of all planes and all assigned points to these planes, we register the points to have a minimal distance to their respective planes.
The transformation $T(\vec{p}_i)$ of each point $\vec{p}_i$ with respect to a 6 DoF motion is described in homogeneous coordinates using the roll-pitch-yaw ($\varphi-\vartheta-\psi$) Tait-Brian angles as in~\cite{diebel2006representing}. Transforming the result back from homogeneous coordinates and using $C_a$ and $S_a$ to denote the cosine and sine of the angle in the subscript, and $t_a$ denoting the translation along the axis in the subscript yields:
\begin{align}
	T(\vec{p}_i)  =
    \resizebox{0.35\textwidth}{!}{$\begin{bmatrix}
        x_i C_\vartheta C_\psi - y_iC_\vartheta S_\psi + z_i S_\vartheta + t_x\\
        x_i (C_\varphi S_\psi + C_\psi S_\varphi S_\vartheta) + y_i(C_\varphi C_\psi - S_\varphi S_\vartheta S_\psi) - z_i C_\vartheta S_\varphi + t_y\\
        x_i(S_\varphi S_\psi - C_\varphi C_\psi S_\vartheta) + y_i(C_\psi S_\varphi + C_\varphi S_\vartheta S_\psi) + z_i C_\varphi C_\vartheta + t_z
    \end{bmatrix}$}
\end{align}
From this we define the function $D(\varphi,\vartheta,\psi,t_x,t_y,t_z, \vec{p}_i)$ that computes the distance of a point $\vec{p}_i$ to its corresponding plane $\mathcal{P}_k$.
Omitting the arguments of the function for simplicity:
\begin{align}
\begin{aligned}
    D &= T(\vec{p}_i) \cdot \vec{n}_{\mathcal{P}_k} \\
      &= n_{\mathcal{P}_k}^x (x_i C_\vartheta C_\psi - y_iC_\vartheta S_\psi + z_i S_\vartheta + t_x)\\
       &+ n_{\mathcal{P}_k}^y(x_i (C_\varphi S_\psi + C_\psi S_\varphi S_\vartheta)\\
       &\qquad+ y_i(C_\varphi C_\psi - S_\varphi S_\vartheta S_\psi) - z_i C_\vartheta S_\varphi + t_y)\\
       &+ n_{\mathcal{P}_k}^z (x_i(S_\varphi S_\psi - C_\varphi C_\psi S_\vartheta)\\
       &\qquad+ y_i(C_\psi S_\varphi + C_\varphi S_\vartheta S_\psi) + z_i C_\varphi C_\vartheta + t_z)\\
       &- \rho_{\mathcal{P}_k}
\end{aligned}
\end{align}

This distance function is what we want to minimize for all points and their respective planes.
Hence the error function $E$ is chosen as the square of the L2-norm of the distance:
\begin{align}
	E = \sum_{\forall \mathcal{P}_k}\sum_{\vec{p}_i \in \mathcal{P}_k} \norm[2]{D(\varphi,\vartheta,\psi,t_x,t_y,t_z, \vec{p}_i)}^2
\end{align}
Its gradient follows then immediately:
\begin{align}
	&\nabla E =  \sum_{\forall \mathcal{P}_k}\sum_{\vec{p}_i \in \mathcal{P}_k} \resizebox{0.28\textwidth}{!}{$\begin{bmatrix}\pardiff{\varphi} E&\pardiff{\vartheta} E&\pardiff{\psi} E&\pardiff{t_x} E&\pardiff{t_y} E&\pardiff{t_z} E\end{bmatrix}^T$}\\
    &\Rightarrow\nabla E = \resizebox{0.35\textwidth}{!}{$\sum_{\forall \mathcal{P}_k} \sum_{\vec{p}_i\in \mathcal{P}_k}2D(\vec{\Pi}, \vec{p}_i)\begin{bmatrix}\nabla E_\varphi &\nabla E_\vartheta & \nabla E_\psi & n_{\mathcal{P}_k}^x&n_{\mathcal{P}_k}^y&n_{\mathcal{P}_k}^z\end{bmatrix}^T$}
\end{align}
Where
\begin{align}
    &\begin{alignedat}{3}
        \nabla E_\varphi = &n_{\mathcal{P}_k}^y&&(x_i[-S_\varphi S_\psi+C_\varphi C_\psi S_\vartheta] \\&&&+y_i[-S_\varphi C_\psi -C_\varphi S_\vartheta S\psi] \\&&&- z_i C_\varphi C_\vartheta)\\
        &+ n_{\mathcal{P}_k}^z&&(x_i[C_\varphi S_\psi+C_\varphi C_\psi S_\vartheta] \\&&&+y_i[C_\varphi C_\psi -S_\varphi S_\vartheta S\psi] \\&&&- z_i S_\varphi C_\vartheta)
    \end{alignedat} \\
    &\begin{alignedat}{3}
        \nabla E_\vartheta = & n_{\mathcal{P}_k}^x(-x_iS_\vartheta C_\psi + y_i S_\vartheta S_\psi + z_iC_\vartheta) \\
        &+ n_{\mathcal{P}_k}^y(x_iC_\psi S_\varphi C_\vartheta - y_iS_\varphi C_\vartheta S_\psi + z_i S_\vartheta S_\varphi)  \\
        &+ n_{\mathcal{P}_i}^z(-x_iC_\varphi C_\psi C_\vartheta + y_iC_\varphi C_\vartheta S_\psi - z_i C_\varphi S_\vartheta)
    \end{alignedat}\\
    &\begin{alignedat}{3}
       \nabla E_\psi =& n_{\mathcal{P}_k}^x&&\left(-x_iC_\vartheta S_\psi - y_iC_\vartheta C_\psi\right) \\
       &+ n_{\mathcal{P}_k}^y&&(x_i[C_\varphi C_\psi - S_\varphi S_\vartheta S_\psi] \\
       &&&+ y_i[-C_\varphi S_\psi - S_\varphi S_\theta C_\psi]) \\
       &+ n_{\mathcal{P}_k}^z&&(x_i[S_\varphi C_\psi + C_\varphi S_\psi S_\theta] \\
       &&&+ y_i[-S_\psi S_ \varphi + C_\varphi S_\vartheta C_\psi]) 
    \end{alignedat}
\end{align}
and $\vec{\Pi}=\begin{bmatrix}\varphi & \vartheta & \psi & t_x & t_y & t_z\end{bmatrix}^T$.

As the gradient is well-defined we minimize the error function with any gradient based method. 
The commonly used, well-known stochastic gradient descent (SDG) algorithm computes 
\begin{align}
    \vec{\Pi}_{k+1} = \vec{\Pi}_{k} - \alpha \nabla E
\end{align}
where $\alpha$ is the learning rate.
To accelerate convergence and to improve the found solution further modifications are made.

@@

Another consideration to speed up SDG is to adaptively recalculate $\boldsymbol\alpha$ for each iteration. 
We employ and modify ADADELTA as a technique to do so, which is described in detail in~\cite{zeiler2012adadelta}.
The main idea is the following:
It extends the SDG algorithm by two terms.
First, an exponentially decaying average of past gradients $\vec{G}_k$, which is recursively defined as
\begin{align}
    \vec{G}_{k+1} = \zeta \vec{G}_{k} + (1 - \zeta) {\nabla E}^2
\end{align}
and second, an exponentially decaying average of past changes $\vec{X}_k$, which is defined as
\begin{align}
    \vec{X}_{k+1} = \zeta \vec{X}_{k} + (1 - \zeta) {\boldsymbol\alpha \nabla E}^2
\end{align}
where $\zeta \leq 1$ is a decay constant, typically close to $1$.
The root mean squared (RMS) of these quantities are
\begin{align}
    RMS[\vec{G}]_{k} = \sqrt{\vec{G}_{k} + \epsilon_{num}}
\end{align}
and 
\begin{align}
    RMS[\vec{X}]_{k} = \sqrt{\vec{X}_{k} + \epsilon_{num}}
\end{align}
where $\epsilon_{num} > 0$ is a very small constant, typically close to $0$ (Note: this is a different threshold as the one used in section~\ref{sec:point-to-plane-correspondences}).
It will prevent dividing by zero in the recalculation of $\boldsymbol\alpha$, which is as follows:
\begin{align}
    \boldsymbol\alpha_{k} = \frac{RMS[X]_{k-1}}{RMS[G]_{k}}
    \label{eq:adadeltaalphaupdate}
\end{align} 
For our particular application, ADADELTA behaves a little too aggressively.
Despite giving a good measure on how to adapt $\boldsymbol\alpha$, the algorithm sometimes overshoots, and does not converge.
Therefore, we employ another scaling factor, typically not found in ADADELTA, extending eq.~\eqref{eq:adadeltaalphaupdate} to:
\begin{align}
    \boldsymbol\alpha_{k} = \boldsymbol\alpha_0 \cdot \frac{RMS[X]_{k-1}}{RMS[G]_{k}}
    \label{eq:adafinalized}
\end{align} 
where $\boldsymbol\alpha_0 $ holds the scaling factors for each dimension.

@@

Firstly, we introduce the possibility to choose a lock on some dimensions from being optimized by setting the corresponding $\alpha_i$ to zero.
Although 6D optimization generally works, reducing the optimization space is particularly useful if the source of error in the system is known and a model exists.
That way, e.g., as we expect a spherical robot to move on a plane the position along the axis perpendicular to the plane is constant and should not be used for optimization.\todo{@Fabi clarify in experiments which locks were active.}

### 2021-06-28 Latex/sections/results.tex
Removed mean as its not representative: Includes points with distance > 100 only with distance 100
Done reviewer 2 additional points corrections

The plane-based registration is applied to the simulated dataset with noisy pose and range measurements (cf. Figure~\ref{fig:simulatedDatasets}) without further processing.
Assuming this represents a coarsely pre-registered 3D point cloud, the distances to the ground truth were evaluated before and after the plane-based registration. 
Figure~\ref{fig:simulatedEvaluation} shows the different point-to-point distances.

\begin{figure*}
 	\centering
 	\begin{minipage}[c]{0.495\textwidth}
 		\centering
		\includegraphics[width=\textwidth]{./images/uncorr_bottom_pose}\\
		\includegraphics[width=\textwidth]{./images/uncorr_side_view}\\
  		\includegraphics[width=\textwidth]{./images/uncorr_hist}
  	\end{minipage}\hfill
  	\begin{minipage}[c]{0.495\textwidth}
  		\centering
		\includegraphics[width=\textwidth]{./images/corr_bottom_pose}\\
		\includegraphics[width=\textwidth]{./images/corr_side_view}\\
  		\includegraphics[width=\textwidth]{./images/corr_hist}
  	\end{minipage} 	
 	\caption{Evaluation of point distances before (left) and after (right) the plane-based registration on a simulated dataset. A maximal distance of \SI{2}{m} is set, such that all points that display a higher distance value are excluded from the analysis. Further, the histogram joins all values with a distance greater than \SI{1}{m} into the last bucket. The top two columns show a heat map of distances, while the bottom shows the corresponding histogram. The color mapping is equivalent in both. An animation of the matching process is given at \url{https://youtu.be/0ps5Pg4qo4E} .}
 	\label{fig:simulatedEvaluation}
\end{figure*}

@@

\begin{table}
	\centering
	\begin{tabular}{@{}lcccc@{}}\hline
		& P90 & P95 & P98 & Mean\\ \hline\hline
		Uncorrected & \SI{24.2}{\centi\meter} &  \SI{38.2}{\centi\meter} &  \SI{61.5}{\centi\meter} & \SI{9.7}{\centi\meter}\\
		Corrected & \SI{12.8}{\centi\meter} &  \SI{16.6}{\centi\meter} &  \SI{21.6}{\centi\meter} & \SI{6.3}{\centi\meter}\\ \hline
	\end{tabular}
	\caption{Comparison of point-distances in the uncorrected and corrected simulated dataset.}
	\label{tab:percentiles}
\end{table}

@@

Figure~\ref{fig:cylon-corrected} shows the results obtained before and after employing the plane based registration on the dataset acquired by the floating sphere.
We combine ten scans into one meta-scan, which is then globally registered.
This does not only speed up convergence due to the proportional effect on the error function but also decreases the risk of transforming the scan in a senseless way.
Transformations like these happen when incorrectly labeled points have a huge effect on the error gradient, which is especially the case when considering only a few points.

>>>>

The plane-based registration is applied to the simulated dataset with noisy pose and range measurements (cf. Figure~\ref{fig:simulatedDatasets}) without further processing.
Assuming this represents a coarsely pre-registered 3D point cloud, the distances to the ground truth were evaluated before and after the plane-based registration. 
Figure~\ref{fig:simulatedEvaluation} \todo{REDO experiments!!!} shows the different point-to-point distances.

\begin{figure*}
 	\centering
 	\begin{minipage}[c]{0.495\textwidth}
 		\centering
		\includegraphics[width=\textwidth]{./images/uncorr_bottom_pose}\\
		\includegraphics[width=\textwidth]{./images/uncorr_side_view}\\
  		\includegraphics[width=\textwidth]{./images/uncorr_hist}
  	\end{minipage}\hfill
  	\begin{minipage}[c]{0.495\textwidth}
  		\centering
		\includegraphics[width=\textwidth]{./images/corr_bottom_pose}\\
		\includegraphics[width=\textwidth]{./images/corr_side_view}\\
  		\includegraphics[width=\textwidth]{./images/corr_hist}
  	\end{minipage} 	
 	\caption{Evaluation of point distances before (left) and after (right) the plane-based registration on a simulated dataset. Lateral images always have the same orientation. A maximal distance of \SI{2}{m} is set, such that all points that display a higher distance value are excluded from the analysis. Further, the histogram joins all values with a distance greater than \SI{1}{m} into the last bucket. The top two columns show a heat map of distances, while the bottom shows the corresponding histogram. The color mapping is equivalent in both. An animation of the matching process is given at \url{https://youtu.be/0ps5Pg4qo4E}.}
 	\label{fig:simulatedEvaluation}
\end{figure*}

@@

\begin{table}
	\centering
	\begin{tabular}{@{}lccc@{}}\hline
		& P90 & P95 & P98 \\ \hline\hline
		Uncorrected & \SI{24.2}{\centi\meter} &  \SI{38.2}{\centi\meter} &  \SI{61.5}{\centi\meter} \\%& \SI{9.7}{\centi\meter}\\
		Corrected & \SI{12.8}{\centi\meter} &  \SI{16.6}{\centi\meter} &  \SI{21.6}{\centi\meter} \\\hline%& \SI{6.3}{\centi\meter}\\ \hline
	\end{tabular}
	\caption{Comparison of point-distances in the uncorrected and corrected simulated dataset.}
	\label{tab:percentiles}
\end{table}

The second significant peak in the right hand histogram in figure~\ref{fig:simulatedEvaluation} \todo{Explain this peak. Not sure right now.}

@@

Figure~\ref{fig:cylon-corrected} shows the results obtained before and after employing the plane based registration on the dataset acquired by the floating sphere.
The pre-registration is obtained by determining the orientation of the sphere via Madgwick filtered~\cite{madgwick2010efficient} IMU measurements.
To increase the number of possible point-to-plane correspondences we combine ten temporally successive scans into one meta-scan, which is then globally registered. 
For this we must assume constant pose over this time duration. 
However, given the high measurement frequency of the Lidar this introduces only a small error. 
This does not only speed up convergence due to the proportional effect on the error function but also decreases the risk of transforming the scan in a senseless way.
Transformations like these happen in particular for a small collection of points as outliers have more influence.

### 2021-06-28 Latex/sections/simulation.tex
Done reviewer 2 additional points corrections

The Livox-Mid100 laser scanner consists of three Mid40 scanners arranged to provide an overall field-of-view of \SI{98.4}{\degree} (horizontal) by \SI{38.4}{\degree} (vertical).
With a range precision of \SI{2}{\centi\meter} up to \SI{20}{\meter} range it provides \SI[per-mode=symbol]{300,000}{\pts\per\second} with a range precision of \SI{2}{\centi\meter} up to \SI{20}{\meter} range.
A particular advantage of the Livox laser scanners is their unique scanning pattern.
The scanners use two motorized, so-called Risley prisms to steer the beam deflectance.
For the specific ratio of rotation speeds used in the Livox scanners, the beam traces out a non-repetitive, flower-shaped scanning pattern~\cite{thorlabs}.
Since the pattern is non-repetitive, the point density increases with integration time, enabling denser measurements.

@@

Accordingly, a spherical robot that is assumed to be translating exactly along one axis by rotating about one other axis at an angular velocity $\omega$  experiences random disturbance torques at each time instant that accelerate the rotation of the sphere about the respective axes. 
This additional motion is determined by integrating the disturbance torques, which are sampled from some normal distribution $n_\phi, n_\psi \in \mathcal{N}(\mu,\sigma^2)$ each.  
The pose update in the discrete simulation is therefore given by: 
\begin{align}
	\begin{bmatrix}\varphi\\\vartheta\\\psi\\x\\y\\z\end{bmatrix}_{k+1} = 
	\begin{bmatrix}\varphi\\\vartheta\\\psi\\x\\y\\z\end{bmatrix}_{k} 
	+ \begin{bmatrix}1\\0\\0\\R\\0\\0\end{bmatrix}\omega \Delta t  + \Delta t \underbrace{\sum_{i = 0}^k \begin{bmatrix}n_\varphi[i] \Delta t\\0\\n_\psi[i]\Delta t\\0\\0\\0\end{bmatrix}}_{\substack{\text{Difference in}\\\text{angular velocity}}}
\end{align}

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.495\textwidth]{images/perfect_bottom}\hfill
	\includegraphics[width=0.495\textwidth]{images/noisy_pose_and_range_bottom}\\
	\includegraphics[width=0.495\textwidth]{images/perfect_top}\hfill
	\includegraphics[width=0.495\textwidth]{images/noisy_pose_and_range_top}\\
	\includegraphics[width=0.495\textwidth]{images/perfect_top_view}\hfill
	\includegraphics[width=0.495\textwidth, height=0.1825\textwidth]{images/noisy_pose_and_range_top_view}
	\caption{Two simulated datasets of a long hallway of size $\SI{4}{\meter}\times\SI{3}{\meter}\times\SI{100}{\meter}$. Ideal noise free (left), noisy pose ($\mathcal{N}_\varphi = \mathcal{N}_\psi(\mu = 0.0001, \sigma= 0.00001)$) and noisy range measurements ($\mathcal{N}_r(\mu = 0, \sigma = 0.001)$) (right).}
	\label{fig:simulatedDatasets}
\end{figure*}

>>>>

The Livox-Mid100 laser scanner consists of three Mid40 scanners arranged to provide an overall field-of-view of \SI{98.4}{\degree} (horizontal) by \SI{38.4}{\degree} (vertical).
It provides \SI[per-mode=symbol]{300,000}{\pts\per\second} with a range precision of \SI{2}{\centi\meter} up to \SI{20}{\meter} range~\cite{LivoxMid40-100}.
A particular advantage of the Livox laser scanners is their unique scanning pattern.
The scanners use two motorized, so-called Risley prisms to steer the beam deflectance.
For the specific ratio of rotation speeds used in the Livox scanners, the beam traces out a non-repetitive, flower-shaped scanning pattern~\cite{thorlabs}.
Since the pattern is non-repetitive, the point density increases with integration time, enabling denser measurements.

@@

Accordingly, a spherical robot that is assumed to be translating exactly along one axis by rotating about one other axis at an angular velocity $\omega$  experiences random disturbance torques at each time instant that accelerate the rotation of the sphere about the respective axes. 
This additional motion is determined by integrating the disturbance torques, which are sampled from some normal distribution $n_\phi, n_\psi \in \mathcal{N}(\mu,\sigma^2)$ each.  
The pose update after a discrete time step $\Delta t$ in the simulation is therefore given by: 
\begin{align}
	\begin{bmatrix}\varphi\\\vartheta\\\psi\\x\\y\\z\end{bmatrix}_{k+1} = 
	\begin{bmatrix}\varphi\\\vartheta\\\psi\\x\\y\\z\end{bmatrix}_{k} 
	+ \begin{bmatrix}1\\0\\0\\R\\0\\0\end{bmatrix}\omega \Delta t  + \Delta t \underbrace{\sum_{i = 0}^k \begin{bmatrix}n_\varphi[i] \Delta t\\0\\n_\psi[i]\Delta t\\0\\0\\0\end{bmatrix}}_{\substack{\text{Difference in}\\\text{angular velocity}}}
\end{align}
where $\Delta t$ is generally chosen as some small value (e.g. \SI{1}{\milli\second})

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.495\textwidth]{images/perfect_bottom}\hfill
	\includegraphics[width=0.495\textwidth]{images/noisy_pose_and_range_bottom}\\
	\includegraphics[width=0.495\textwidth]{images/perfect_top}\hfill
	\includegraphics[width=0.495\textwidth]{images/noisy_pose_and_range_top}\\
	\includegraphics[width=0.495\textwidth]{images/perfect_top_view}\hfill
	\includegraphics[width=0.495\textwidth, height=0.1825\textwidth]{images/noisy_pose_and_range_top_view}
	\caption{Two simulated datasets of a long hallway of size $\SI{4}{\meter}\times\SI{3}{\meter}\times\SI{100}{\meter}$. Ideal noise free (left), noisy pose ($\mathcal{N}_\varphi = \mathcal{N}_\psi(\mu = 0.0001, \sigma= 0.00001)$) and noisy range measurements ($\mathcal{N}_r(\mu = 0, \sigma = 0.001)$) (right). 
	The normal distributions are define in such a way, that in the resulting dataset plane detection is still possible.}
	\label{fig:simulatedDatasets}
\end{figure*}

### 2021-06-28 Latex/sections/stateOfTheArt.tex
Done reviewer 2 additional points corrections

Both approaches use plane-to-plane correspondences to pre-register the scans.
However, for pre-registration the classical point-to-point registration is also very effective. 
One advantage that point-to-point correspondences have over plane-to-plane correspondences is that they do not require a long stop in each pose.
For plane-to-plane correspondences, this is necessary to gather enough data to measure planes in each scan robustly.
The resulting scan procedure is stop-scan-and-go.
In particular, for the application of a spherical robot this standstill in each pose cannot be guaranteed or even approximated, making continuous-time approaches using point-to-plane correspondences the method of choice.

>>>>

Both approaches use plane-to-plane correspondences to pre-register the scans.
However, for pre-registration the classical point-to-point registration is also very effective. 
One advantage that point-to-point correspondences have over plane-to-plane correspondences is that they do not require a long stop in order to obtain enough points to detect planes in each pose.
For plane-to-plane correspondences, this is necessary to gather enough data to measure planes in each scan robustly.
The resulting scan procedure is stop-scan-and-go.
In particular, for the application of a spherical robot this standstill in each pose cannot be guaranteed or even approximated, making continuous-time approaches using point-to-plane correspondences the method of choice.
